---
title: 'Supplementary materials: Examples of entropy calculations with R'
author: "Edgar Schach, Raimon Tolosana et al."
date: "2023-07-12"
output:
  html_document:
    df_print: paged
---

## System setup

This document is an R-markdown file for illustration of the main findings of the paper xxxx by xxxx. R-markdown files can be interactively edited, executed and visualised with the integrated developing interface [R-Studio](www.posit.com/rstudio) for the data science [R programming language](www.r-project.org)

```{r} 
# loading packages
library(tidyverse)    # set of packages for data science
library(purrr)        # package for parallel data manipulation
library(RColorBrewer) # color scales
library(ggpubr)       # publication-ready diagrams

# define entropy colors and entropy contribution sources 
cols_component.dist <- brewer.pal(n = 5, name = "Blues")[2:5]
names(cols_component.dist) <- c("c0j0","cij0","c0jk","cijk")

cols_particle.dist <- brewer.pal(n = 4, name = "Greens")
names(cols_particle.dist) <- c("p00k","p0jk","pi0k","pijk")

cols_partition.vec <- brewer.pal(n = 4, name = "Reds")
names(cols_partition.vec) <- c("ti00","ti0k","tij0","tijk")

entropy_colors <- c(cols_component.dist,
                    cols_particle.dist,
                    cols_partition.vec)

type.levels <- c("ti00","ti0k","tij0","tijk", # partitions
                 "p00k","p0jk","pi0k","pijk", # distributions
                 "c0j0","cij0","c0jk","cijk") # compositions

# load computing functions 
source("Entropy_Functions-toSubmit.R")
```


## Creating Particles

The function `particles_create()` generates a particle data set, with $K$ particles and $J$ components:
```{r}
set.seed(523)
feed_part <- particle_create(K = 100, J =2, l = 0.2)
head(feed_part) # show a few lines of the particle object
```

This can be plotted as a stacked bar plot, with bar total height indicating particle size: 
```{r}
feed_part_df <- data.frame(feed_part, particle = 1:nrow(feed_part) ) %>% gather(key = "Components", value = "mass", -particle) 

g <- ggplot(feed_part_df,aes(x = particle, y = mass, fill = Components)) +
  geom_bar(stat = "identity") +
  theme_bw() + ylab("size")
print(g)
ggsave(filename= "feed.part.png",device = "png", units = "cm",dpi = 300,width = 15,height = 7.5)
```



## On the comminution models

Given the 0-dimensional description of particles, comminution simulation is simulated as well at that conceptual level. The following two breakage styles are considered:

- **Liberation breakage** generates, out of each particle, at most $J$ child particles, each containing the whole mass of one of the components of the parent particle. Already liberated particles do not break further.
- **Random breakage** generates, for each parent particle, $L$ child particles ($L$ equally probable integer between 2 and 5), each capturing a random proportion $\boldsymbol{\pi}_j$ of the mass of each of the $J$ components present. These random proportions are drawn from a symmetric Dirichlet distribution $\mathcal{D}i(a\cdot \mathbf{1})$ with a a real value $a$ uniformly drawn from the interval $(0.01, 5)$. Each particle breaks with different values of $\{L, \boldsymbol{\pi}_1, \boldsymbol{\pi}_2, \ldots, \boldsymbol{\pi}_J, a\}$.  

These two modes are combined in a global `comminute()` function, which can control the proportion of breaking particles subject to random breakage, and the proportion of particles being broken (in which case, larger particles tend to break faster than smaller ones). 

The following shows four settings, with particles breaking with different proportions of random breakage

```{r}
# break feed following several styles
set.seed(523)
perfect.liberation = comminute(feed_part, prob_rnd_break = 0)
overgrinding = comminute(perfect.liberation$progeny)
set.seed(523)
realistic.comminution = comminute(feed_part, prob_rnd_break = 0.5)
set.seed(523)
random.breakage = comminute(feed_part, prob_rnd_break = 1)

# combine in a list
comminution.res <- list(
  perfect.liberation = perfect.liberation,
  overgrinding = overgrinding,
  realistic.comminution = realistic.comminution,
  random.breakage = random.breakage
)

# define a function for plotting one breakage product
doOnePlot <- function(partset, title){
    partset_df <- data.frame(partset$progeny, particle = 1:nrow(partset$progeny) ) %>% 
      gather(key = "Components", value = "mass", -particle) 

 g <- ggplot(partset_df, aes(x = particle, y = mass, fill = Components)) +
  geom_bar(stat = "identity") + ylab("size") +
  theme_bw() + ggtitle(title) 
 return(g)
}

# compute plots for all breakage products
plots <- map2(comminution.res, 
              c("perfect liberation",  "overgrinding", 
                "realistic comminution", "random breakage"), 
              doOnePlot)

# combine
ggpubr::ggarrange(plotlist = plots, 
          nrow = 2, ncol=2,
          common.legend = T,
          legend = "bottom")
```


## On the partition curve model

We use for our calculations Rogers' model for density separation, programmed in the function `separate()`. This function has a sharpness parameter `alpha`, which can take three kinds of values:

- for any real value of `alpha`, this is taken as the $alpha$ parameter of the Rogers' partition model, 

$$
t_c(\rho; a, \alpha, \rho_0) = a + (1-a)\frac{1}{1+\frac{\rho_0}{\rho}\exp\left[\alpha\left(1-\frac{\rho^3}{\rho_0^3}\right)\right]} 
$$
where $\rho_0$ represents the cutoff density and $a$ is a maximum recovery;
- for `alpha` infinite, a perfect separation model is taken, 

$$
t_c(\rho; \alpha, \rho_0) = 
\left\{
\begin{array}{cl}
 1 & \text{if  } \mathrm{sign}(\alpha)\rho >\mathrm{sign}(\alpha)\rho_0 \\
 0 & \text{otherwise}
\end{array}
\right.
$$
- for `alpha=NA` (not available), a perfect splitting is considered, e.g. with $t_c(\rho) = 0.5$. 

The following code generates examples for all cases:
```{r}
# set of alpha values to consider
alpha <- c(NA, 0:10, Inf)

# densities specified for each phase
densities <- c(A = 2.7, B = 7) # e.g. quartz and cassiterite

# construct names out of the alpha levels considered
nms <- sapply(alpha,function(x) paste("alpha =",x,sep = " ") )
# compute and extract partition coefficient 
res <- lapply(alpha, separate, cut.point = 5, phase.property = densities, feed = feed_part)
res <- lapply(res, function(x) data.frame(density = x$prop.vec, partition = x$prob.vec))
names(res) <- nms

# change alpha names to explicit mention of perfect models
nb.cols <- length(res)
names(res) = c("perfect splitting","alpha = 0","alpha = 1","alpha = 2","alpha = 3","alpha = 4","alpha = 5","alpha = 6","alpha = 7","alpha = 8","alpha = 9","alpha = 10","perfect separation")
nms <- names(res)

# recast results as data set and plot
res_df <- bind_rows(res,.id = "alpha")
res_df$alpha <- factor(res_df$alpha, ordered=TRUE, levels=nms, labels=nms)

mycolors <- colorRampPalette(brewer.pal(8, "Spectral"))(nb.cols)

g <- ggplot(res_df,aes(x = density,y = partition, col = alpha))+
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(x = "particle density in g/cm³",y = "partition", col = "alpha levels") +
  scale_color_manual(values = mycolors)
print(g)
ggsave(filename= "rogers.function.png",device = "png", units = "cm",dpi = 300,width = 15,height = 10)
```


## Entropy of comminution products

With the comminution products simulated before, we can now compute batch entropies and compare the effect of the several comminution modes on entropy contributions:

```{r}
# compute entropy of each batch, inlcuding feed
comminution.res <- lapply(comminution.res,function(x) x$progeny)
comminution.res[["feed"]] <- feed_part
comminution.res_ent <- lapply(comminution.res, compute_batch_entropy)
comminution.res_ent <- bind_rows(comminution.res_ent,.id = "scenario")

# change names
comminution.res_ent$scenario <-
  fct_relevel(comminution.res_ent$scenario,
              "feed", "perfect.liberation", "overgrinding",
              "realistic.comminution", "random.breakage")

labs_scenarios <- 
  c("feed", "perfect liberation", "overgrinding", "reference comminution", "random breakage")
names(labs_scenarios) <- levels(comminution.res_ent$scenario)

# plot!
g <- ggplot(comminution.res_ent,aes(x = .id, y = value, fill = factor(contribution, levels = type.levels))) +
  geom_bar(stat = "identity") +
  facet_grid(.~scenario,scales = "free", space = "free", labeller = labeller(scenario = labs_scenarios))+
  theme_bw() +
 scale_fill_manual(values = entropy_colors) +
  labs(x = " ", y = expression(Delta~ ~s*"*"), fill = "contribution")
print(g)
ggsave(filename= "component.dispersion.png",device = "png", units = "cm",dpi = 300,width = 21,height = 10)
```


## Evolution of entropy through comminution and separation 

The following code combines realistic setups for comminution and separation, and computes the entropy of the several stages:

```{r echo=FALSE}
#milling feed particles
set.seed(523)
milling.prod <- comminute(as.matrix(feed_part), prob_rnd_break = 0.5)
milling.prod <- milling.prod$progeny

#separation of milled feed
set.seed(11523)
res_sep <- separate(feed = milling.prod, cut.point = 5, 
                    phase.property = densities, alpha = 5)

#computing entropies for different stages
res <- list(feed = compute_batch_entropy(feed_part),
            comminution = compute_batch_entropy(milling.prod),
            separation = compute_stage_entropy(res_sep))
# mass_pulls = attr(res$separation, "mass_pulls")
res <- bind_rows(res,.id = "stage")

#plotting entropies
g <- ggplot(res,aes(x = .id, y = value, fill = factor(contribution,levels = type.levels))) +
  geom_bar(stat = "identity") +
  facet_grid(.~fct_relevel(stage,"feed"),scales = "free", space = "free")+
  theme_bw() +
 scale_fill_manual(values = entropy_colors) +
  labs(x = "stage",y =expression(Delta~ ~s*"*"), fill = "contribution")
print(g)

res_feed <- res %>% filter(stage == "feed")
res_comminution <- res %>% filter(stage == "comminution")
res_separation <- res %>% filter(stage == "separation")
ent.max <- res %>% group_by(stage,.id) %>% summarize(sum_ent = sum(value))
ent.max <- max(ent.max$sum_ent)
lims <- c(0,ent.max)

g <- ggplot(res_feed,aes(x = .id, y = value, fill = factor(contribution,levels = type.levels))) +
  geom_bar(stat = "identity") +
  theme_bw() +
 scale_fill_manual(values = entropy_colors) +
 scale_y_continuous(limits = lims,oob = scales::oob_keep) +
  labs(title = "feed",x = "",y = expression(Delta~ ~s*"*"), fill = "contribution")
print(g)

h <- ggplot(res_comminution, aes(x = .id, y = value, fill = factor(contribution, levels = type.levels))) +
  geom_bar(stat = "identity") +
  theme_bw() +
 scale_fill_manual(values = entropy_colors) +
 scale_y_continuous(limits = lims,oob = scales::oob_keep) +
  labs(title = "comminution",x = "",y = expression(Delta~ ~s*"*"), fill = "contribution")
print(h)



i <- ggplot(res_separation,aes(x = .id, y = value, fill = factor(contribution, levels = type.levels))) +
  geom_bar(stat = "identity") +
  theme_bw() +
 scale_fill_manual(values = entropy_colors) +
 scale_y_continuous(limits = lims, oob = scales::oob_keep) +
  labs(title = " ",x = "",y = expression(Delta~ ~s*"*"), fill = "contribution") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.5, size = 12))
print(i)
ggsave(filename= "entropy.aggregations.png",device = "png", units = "cm",dpi = 300,width = 15,height = 10)

dat <- res_separation %>% filter(.id %in% c("ti00.cij0.pijk","ti00.pi0k.cijk"))

j <- ggplot(dat,aes(x = .id, y = value, fill = factor(contribution, levels = type.levels))) +
  geom_bar(stat = "identity") +
  theme_bw() +
 scale_fill_manual(values = entropy_colors) +
 scale_y_continuous(limits = lims,oob = scales::oob_keep) +
  labs(title = "separation",x = "",y = expression(Delta~ ~s*"*"), fill = "contribution")
print(i)



plots = list(feed = g,
             comminution = h,
             separation = j)

ggarrange(plotlist = plots,nrow = 1,
          common.legend = T,
          legend = "right",
          hjust = c(-2,-0.4,-1),
          vjust = 0.5,
          heights = c(0.7,0.7,0.7),
         legend.grob = get_legend(plots),
          widths = c(1,1,1))


ggsave(filename= "flowsheet.png",device = "png", units = "cm",dpi = 300,width = 20,height = 10)

```


## Entropy generation by separation

This code explores the role of separation on increasing the uncertainty about the fate of particles, and how this uncertainty is partly or totally compensated by the generation of purer products:

```{r}
# separation model parameters
densities <- c(2.7,7)
cut.point <- 5

# compute separations
set.seed(11523)
sep_res <- 
  list(perfect.separation=Inf, realistic.separation=5, perfect.splitting=NA) %>%
  lapply( function(alpha){
    separate(feed = realistic.comminution$progeny, 
             phase.property = densities, 
             alpha=alpha, cut.point = 5)
  } )

# compute entropies
sep_res_ent <- lapply(sep_res, compute_stage_entropy)

# add initial entropy of the feed
sep_res_ent[["feed"]] <- compute_batch_entropy(realistic.comminution$progeny)

# reform outcome for plotting
sep_res_ent <- bind_rows(sep_res_ent,.id = "scenario")
a <- sep_res_ent$.id
a[a == "c0j0.p0jk"] <- "c0j0.p0jk.tijk"
a[a == "p00k.c0jk"] <- "p00k.c0jk.tijk"
sep_res_ent$.id <- a

dat <- sep_res_ent %>% filter(.id %in% c("ti00.cij0.pijk","c0j0.p0jk.tijk"))
dat$scenario <- str_replace(dat$scenario,pattern = "\\."," ")
# plot!
labels_entropy <- c("c0j0.p0jk.tijk", "ti00.cij0.pijk")

g <- ggplot(dat,aes(x = fct_relevel(scenario,"feed","perfect separation","realistic separation","perfect splitting"), y = value, fill = factor(contribution, levels = type.levels))) +
  geom_bar(stat = "identity") +
  facet_grid(.~factor(.id, levels = labels_entropy), scales = "free", space = "free")+
  theme_bw() +
  scale_fill_manual(values = entropy_colors) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.5, size = 12)) +
  labs(x = " ",y = expression(Delta~ ~s*"*"), fill = "contribution") 
 # scale_x_discrete(labels =c("feed","perfect separation","reference separation","perfect splitting"))
print(g)
ggsave(filename= "separation.integrowth.png",device = "png", units = "cm",dpi = 300,width = 15,height = 10)
```

## Optimization of comminution-separation setup

Finally the next code shows how to compute several scenarios for partial comminution followed by a realistic separation, in order to evaluate the best comminution configuration in terms of maximal entropy reduction. Given the inherent variability induced by the simulations, the calculations are repeated 50 times to obtain an expected behaviour:

```{r}
# function doing one calculation
one_simulation = function(nsteps){
 # make space for all consecutive comminution products
 rescom <- matrix(0:nsteps, nrow=1) %>% as.data.frame %>% as.list
 rescom[[1]] =  particle_create(K = 200, J =5, l = 0.05) # feed

 # consecutively mill in nsteps, store only progeny particles
 for(i in 2:(nsteps+1)){
  rescom[[i]] = comminute(rescom[[i-1]], 
                       prob_rnd_break = 0.25, 
                       process_fraction = 0.05 )$progeny
 }

 # apply separation to each milling stage
 densities <- c(2,3,2.5,4.5,6)
 ressep <- lapply(rescom, separate, alpha = 5, phase.property = densities, cut.point = 5)

 # compute entropies post-separation
 ent_com <- lapply(ressep, compute_stage_entropy)

  # extract mass pulls and compute grades and recoveries
 mass_pulls = lapply(ent_com, function(x)attr(x, "mass_pulls") )
 
 # reorder for plotting
 entropies = map_dbl(ent_com, function(x) x %>% filter(.id=="ti00.cij0.pijk") %>% 
  filter(contribution!="ti00") %>% summarize(total = sum(value)) %>% 
    unclass %>% unlist
 )
 
 # compute initial entropy
 ent.feed <- compute_batch_entropy(rescom[[1]])
 ent.feed <- ent.feed %>% filter(.id == "p00k.c0jk") %>% summarise(s = sum(value))
 
 out = 100*(ent.feed$s - entropies)/ent.feed$s

 # attach mass pulls
 attr(out, "mass_pulls") = mass_pulls
 
 # return entropy decreases
 return(out)
}

# compute 50 scenarios
nsteps= 30
set.seed(523)
outcomes_l = replicate(n = 50, expr = one_simulation(nsteps), simplify = FALSE)

# re-form
outcomes = do.call("cbind", outcomes_l)
colnames(outcomes) = paste0("V",1:ncol(outcomes))

# compute average over the scenarios
av_outcome = rowMeans(outcomes)

# plot!
g <- outcomes %>% as.data.frame %>% cbind(cycles=0:nsteps) %>% 
   tidyr::pivot_longer(cols=contains("V"), values_to="entropy") %>%
    ggplot(aes(x = cycles, y = entropy)) +
    geom_path(aes(group=name), color="darkgrey") + 
    geom_path(data=data.frame(cycles = 0:nsteps, entropy=av_outcome), 
              color="red", lwd=2) +   theme_bw() +
  ylab(expression(Delta~ ~s*"*  decrease [%]"))
    
print(g)
ggsave(filename= "minimize.entropy.png",device = "png", units = "cm",dpi = 300,width = 15, height = 10)
```

Additionally, the following diagrams show the evolution of grade and recovery of value mineral (mineral E) on the concentrate (high density product), as a function of the number of grinding cycles. Again, curves for individual simulations are provided in grey, and their average is plotted as a red line.

```{r}
grad_rec = lapply(outcomes_l, 
 function(y){
  mass_pulls = attr(y, "mass_pulls")
  grades = sapply(mass_pulls, function(x) sweep(x, 1, rowSums(x), "/")[1,"E"]) 
  recoveries = sapply(mass_pulls, function(x) sweep(x, 2, colSums(x), "/")[1,"E"]) 
  return(list(grades=grades, recoveries=recoveries))
})

grades = sapply(grad_rec, function(x) x$grades)
recoveries = sapply(grad_rec, function(x) x$recoveries)

colnames(grades) <- colnames(recoveries) <- colnames(outcomes)

av_grade = rowMeans(grades)
av_recovery = rowMeans(recoveries)


# plot!
g_g <- grades %>% as.data.frame %>% cbind(cycles=0:nsteps) %>% 
   tidyr::pivot_longer(cols=contains("V"), values_to="grade") %>%
    ggplot(aes(x = cycles, y = grade)) +
    geom_path(aes(group=name), color="darkgrey") + 
    geom_path(data=data.frame(cycles = 0:nsteps, grade=av_grade), 
              color="red", lwd=2) +   theme_bw() + ylab("grade")

g_r <- recoveries %>% as.data.frame %>% cbind(cycles=0:nsteps) %>% 
   tidyr::pivot_longer(cols=contains("V"), values_to="recovery") %>%
    ggplot(aes(x = cycles, y = recovery)) +
    geom_path(aes(group=name), color="darkgrey") + 
    geom_path(data=data.frame(cycles = 0:nsteps, recovery=av_recovery), 
              color="red", lwd=2) +   theme_bw() + ylab("recovery")

g_gr = egg::ggarrange(g_g, g_r,nrow = 1, ncol=2)
print(g_gr)
```

